  ---
title: "Analyizing data using backward-selection with p-value as the selection criterion and verifying result on diagnostic plots"
author: "Rohan Mate"
date: "November 17, 2016"
output: html_document
---

```{r}
library(lattice)
evals = read.csv("evals.csv", header = TRUE)
```

#1. Describe the distribution of score. Is the distribution skewed? What does that tell you about how students rate courses? Is this what you expected to see? Why, or why not?

```{r}
#The distribution of score:
summary(evals$score)
hist(evals$score)  
-
#Histrograms shows that the score data is skewed to left, which predicts the student are likely to rate course highly.Yes, it was expected to see because the histogram depicts the same as well as median. We expected a normal distribution where most teachers would be rated as average and fewer teachers will be evaluated in the extremes - excellent or unsatisfactory
```

#2. Excluding score, select two other variables and describe their relationship using an appropriate visualization (scatterplot or side-by-side boxplots).

```{r}
#For scatterplot 
plot(evals$age, evals$bty_avg)

#For side by side boxplot
boxplot(evals$age ~ evals$gender)

#Analyzing Scatterplot, we can conclude that as age increases beauty decreases.The variable age is inversly proportional to variable bty_avg.

#Analzing side by side boxplot, there seems no relationship between teachers age and beauty score. It was suspected an inverse relationship 
```
### Multiple Linear Regression

```{r}
plot(evals$bty_avg ~ evals$bty_f1lower )
cor(evals$bty_avg , evals$bty_f1lower )

plot(evals [, 13 : 19 ])
```

#3. Is bty_avg still a significant predictor of score? Has the addition of gender to the model changed the parameter estimate for bty_avg?

```{r}
#Yes, bty_avg is still a significant predictor of score

summary(lm(evals$score ~ evals$bty_avg))
#Adding gender to the equation now enhances the predictor bty_avg as it results in beauty scores with respect to the gender..
```

#4. What is the equation of the line corresponding to males? (Hint: For males, the parameter estimate is multiplied by 1.) For two professors who received the same beauty rating, which gender tends to have the higher course evaluation score?

```{r}
#Equation of the line corresponding to males can be calculated through below equation.
#score^=??^0+??^1×bty_avg+??^2×(1)
#      =??^0+??^1×bty_avg+??^2

#Male professor tends to have 0.17 higher course evaluation score than female professor. Wherein the male professor tends to have the higher course evaluation score for beauty rating

#The action to call, indicator variable "gendermale" rather than "genderfemale" is becuase R simply codes categories alphabetically as 00.(we can change the reference level of a categorical variable, which is the level that is coded as a 0, using therelevel function.)
```

#5. Create a new model called m_bty_rank with gender removed and rank added in. How does R appear to handle categorical variables that have more than two levels? Note that the rank variable has three levels: teaching, tenure track, tenured.

```{r}
m_bty_rank = lm(evals$score ~ evals$bty_avg + evals$rank)
summary(m_bty_rank)

#As rank has three levels (tenure track,tenured,teaching ) two indicator variables are created: one for tenure track and and one for tenured. Teaching is the reference level hence it doesn't show up in the regression output.

#additionally, calculation for bty_avg reflects how much higher a group of professors is expected to score if they have a beauty rating that is one point higher while holding all other variables constant. In above case, it indicates only professors of the same rank with bty_avg scores that are one point apart.
```

#6. The interpretation of the coefficients in multiple regression is slightly different from that of simple regression. The estimate for bty_avg reflects how much higher a group of professors is expected to score if they have a beauty rating that is one point higher w hile holding all other variables constant. In this case, that translates into considering only professors of the same rank with bty_avg scores that are one point apart

```{r}
# Through analysis we indicate, a professor with 1 point more in score than other professors tend to have beauty average of about 0.067 more.

# Wherein, model predicts that, the professors with rank "tenure track" have 0.16 points lesser in score and the professors with rank "tenured" have 0.12 points less in score.For professor with 0 beauty average consist of score of 3.98 points(on avg).

```

#7. Which variable would you expect to have the highest p-value in this model? Why? Hint: Think about which variable would you expect to not have any association with the professor score

```{r}
#I suspect "cls_prof" to have least relationship with score having highest p value.

m_full <- lm ( score ~ rank + ethnicity + gender + language + age + cls_perc_eval + cls_students + cls_level + cls_profs + cls_credits + bty_avg + pic_outfit + pic_color , data = evals )

summary ( m_full )
max_p = max(summary(m_full)$coefficients[, 4])
max_p

which(summary(m_full)$coefficients[, 4] == max_p)

```

#8. Check your suspicions from the previous exercise. Include the model output in your response.

```{r}
plot(evals$score ~ evals$cls_profs)

#As suspected, "cls_professor" has the least relationship with "score".It also has the highest p-value.
```

#9. Interpret the coefficient associated with the ethnicity variable.

```{r}
#The ethnicity of p value for 0.11 describes weak or less relationship to scores and may be dropped as part of the model
```

#10. Drop the variable with the highest p-value and refit the model. Did the coefficients and significance of the other explanatory variables change? (One of the things that makes multiple regression interesting is that coefficient estimates depend on the other variables that are included in the model.) If not,what does this say about whether or not the dropped variable was collinear with the other explanatory variables?

```{r} 
m_full1 = lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval + 
    cls_students + cls_level + cls_credits + bty_avg + pic_outfit + pic_color, 
    data = evals)

summary(m_full1)

#Yes,Removing cls_profs we can identify slight change in coefficients and significance of other explanatory variables. There is more significance now to the level than before indicates all values are slightly lower
```

11. Using backward-selection and p-value as the selection criterion, determine the best model. You do not need to show all steps in your answer, just the output for the final model. Also, write out the linear model for predicting score based on the final model you settle on

```{r} 
m_full2<- lm(score ~ ethnicity + gender + language + age + cls_perc_eval +     cls_credits + bty_avg + pic_color, data = evals)

summary(m_full2)
# Using backward-selection we eleminate all the variables with low p-value.
#linear model:
#x = 3.77 + 0.2 * gender + 0.16 * ethnicity - 0.2 * language - 0.006 * age + 0.004 *cls_perc_eval + 0.5 * cls_credit + 0.05 * bty_avg - 0.19 *pic_color
``` 

#12. Verify that the conditions for this model are reasonable using diagnostic plots.      
```{r}
#To verify conditions for this model are reasonable using diagnostic plots is as follows:

#Each variable is linearly related to the outcome
plot(evals$score ~ evals$ethnicity)
plot(evals$score ~ evals$gender)
plot(evals$score ~ evals$language)
plot(evals$score ~ evals$age)
plot(evals$score ~ evals$cls_perc_eval)
plot(evals$score ~ evals$cls_credits)
plot(evals$score ~ evals$bty_avg)
plot(evals$score ~ evals$pic_color)
#The variables above are linearly related to the score, few are more than others.

```


```{r}

model111=step(lm (evals$score ~ evals$gender + evals$ethnicity + evals$language + evals$age + evals$cls_perc_eval + evals$cls_credits + evals$bty_avg + evals$pic_color, data = "evals",direction="backward"))
```
